A Component-Based Sensor Network for \\ Environmental Monitoring

### 2009-03-14 docs/icccn2009/paper.tex
Cleanup architecture section
git-svn-id: svn://localhost/netbeams/trunk@389 559edde2-a4d4-44e0-96fc-4035e40d7396

By implementing the above Java interface, the resulting code
effectively becomes a DSP Component that can be deployed on a DSP at
runtime. A component is identified by its type (line 3). Only one
component of a certain type per DSP is permissible. A type identifier
must be unique across all DSP Components in a DSP.  Upon deploying a
component, an initialization function is called (line 4). Once a
component is initialized, it may be started (line 4) and stopped (line
7) multiple times. Whenever the DSP wishes to pass a message (that was
generated by some DP), it invokes the \texttt{deliver()} method (line
8). DP sends messages through \texttt{Context} implementation (provided 
by the DSP during DSP initialization)The message structure will be explained in the next section.
One important feature of the DSP is the ability to redeploy itself (maybe
with a new version with updates) at runtime. This allows DSPs to be updated
remotely. This process is managed through messages sent to a special
DSP Component that is able to manage its local DSP.

@@

The basic message structure is defined by an \texttt{AbstractMessage}
type, whose structure serves as a base schema for all different types
of messages. Any message will contain two basic attributes:
\texttt{messageID} and \texttt{ContentType}.  The former uniquely
identifies a message and the latter the data in the body of the
message. Complementary identification information, such as the message
producer and consumer, is located in the header of the message and 
is defined by the schema type \texttt{ComponenIdentifier}. A
\texttt{ComponentIdentifier} carries out information about the
component name and its physical host address.  Finally, the body of
the message represents the payload and may contain any data type
defined by the component developer.

Although XML schema is the de-facto standard to bind and validate XML
data to its type, the architectural infrastructure of DSP uses an XML
data binding technology \cite{xml-dbind} called JAXB \cite{xml-jaxb}.
It is used to generate a type safe API for handling in-memory object
graphs from the XML instances of DSP Messages and the content of its
data payload. Moreover, taking advantage of the data type polymorphism
of XML schema data types -- a concept borrowed from object oriented
principles -- we defined different categories of messages in order to
reflect the nature of the message content. Among those categories are
query and update messages, where the former is used for configuring various
properties supported by DSP Components, and the latter is used for measurement
messages that actually contain sensor readings.

@@

\subsection{DSP Broker}
Broker is responsible for routing the message to one or several DCs. It
distinguishes between local and remote destinations:

\begin{itemize}
\item Local delivery of messages: it uses local object reference to
  deliver the DSP message.
\item Remote delivery of messages: it uses a DSP Component marked as a
  gateway (configured in matching rules) to handle remote message delivery.
\end{itemize}

Broker provides the infrastructure necessary to deliver messages concurrently and
efficiently. Itis  also responsible for setting any additional message header 
information necessary for proper routing. Broker does not dictates which DCs should
receive a message. It delegates this responsibility to the Matcher.

\subsection{Matcher}

Matcher is the component in DSP responsible for matching messages from a DP to one or more
DCs. The decision process is based on evaluation of rules (see Figure
\ref{FIG_MATCHER_RULE}). Rules can be statically or dynamically
configured.

@@

\begin{itemize}
\item The \emph{match criteria} defines the name of the component type
  of the producer and its physical location (IP address);
\item The \emph{match target} defines the name of the component type
  of the consumer, and optionally defines a gateway component.
\end{itemize}

Matcher's main responsibility is to shield DPs from the responsibility
for specifying consumers of their messages. DPs may create messages without any pre-determinted
DCs. Once the message is sent through DSP, the matcher will determine which DCs should
receive the message. Note that, a DP can still specify a DC for its message. In this case, Matcher
includes this DC in list of DCs to receive the message. Matcher also plays an important role on
the dynamic aspect of the DSP. Suppose the communication between two DSPs becomes unreliable and
it necessary to persist messages before they are sent through the communication channel. It is possible to
achieve this goal by deploying a DSP component which implements persistence and modifying the matching
rules to divert messages to the new DSP Component and from the component to the original destination.

>>>>

By implementing the above Java interface, the resulting code
effectively becomes a DSP Component that can be deployed on a DSP at
runtime. A component is identified by its type (line 3). Only one
component of a certain type per DSP is permissible. A type identifier
must be unique across all DSP Components in a DSP. Upon deploying a
component, an initialization function is called (line 4). Once a
component is initialized, it may be started (line 6) and stopped (line
7) multiple times. Whenever the DSP wishes to pass a message (that was
generated by some DP), it invokes the \texttt{deliver()} method (line
8). A DP can send messages via a \texttt{DSPContext} object that was
passed to the DP during initialization.

@@

The basic message structure is defined by an \texttt{AbstractMessage}
type, whose structure serves as a base schema for all different types
of messages. Any message will contain two basic attributes:
\texttt{messageID} and \texttt{ContentType}.  The former uniquely
identifies a message and the latter the data in the body of the
message. Complementary identification information, such as the message
producer and consumer, is located in the header of the message and 
is defined by the schema type \texttt{ComponentIdentifier}. A
\texttt{ComponentIdentifier} carries out information about the
component name and its physical host address.  Finally, the body of
the message represents the payload and may contain any data type
defined by the component developer.

XML schema is the de-facto standard to bind and validate XML
data to its type. Within the DSP we use an XML
data binding technology \cite{xml-dbind} called JAXB \cite{xml-jaxb}.
It is used to generate a type safe API for handling in-memory object
graphs from the XML instances of DSP Messages and the content of its
data payload. Moreover, taking advantage of the data type polymorphism
of XML schema data types, we defined different categories of messages in order to
reflect the nature of the message content. Among those categories are
query and update messages, where the former is used for configuring various
properties supported by DSP Components, and the latter is used for measurement
messages that actually contain sensor readings.

@@

\subsection{DSP Broker and Matcher}

The \emph{broker} is responsible for routing the message to one or
several DCs. It distinguishes between local and remote destinations:

\begin{itemize}
\item Local delivery of messages: it uses local object reference to
  deliver the DSP message.
\item Remote delivery of messages: it uses a DSP Component marked as a
  gateway to handle remote message delivery.
\end{itemize}

The broker is can deliver messages concurrently and asynchronously. It
sets any additional message header information necessary for proper
routing. The broker does not make and routing decisions. It delegates
this responsibility to the \emph{matcher}. The matcher is responsible
for matching messages sent by a DP and route it to one or more DCs.
This decision is based on a set of routing rules (see Figure
\ref{FIG_MATCHER_RULE}) which can be statically or dynamically
configured.

@@

\begin{itemize}
\item The \emph{match criteria} defines the name of the component type
  of the producer and its physical location (IP address).
\item The \emph{match target} defines the name of the component type
  of the consumer, and optionally defines a gateway component.
\end{itemize}

The matcher's main purpose is to achieve a loosely coupling between
DCs and DPs. DPs may create messages without specifying a specific
destination DC. Once the message is sent through DSP, the matcher will
determine which DCs should receive the message. Note that, a DP can
still specify a DC for its message. In this case, the matcher includes
this DC in list of DCs to receive the message.
